Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-21 paper/paper.tex
spaces around tabular
case study in progress

\begin{tabular}{l l}
    \textbf{old} & \textbf{new} \\ \hline
    {}\texttt{[C1 0]} & \texttt{[C1 0]} \\
    {}\texttt{[C2 0]} & \texttt{[C2 0]} \\
    {}\texttt{[C2 1]} & \textit{unused} \\
    {}\texttt{[C1 16]} & \texttt{[C1 16]} \\
    {}\texttt{[C1 17]} & \texttt{[C1 17]} \\
    {}\texttt{[C1 18]} & \textit{unused} \\
    {}\texttt{[C2 16]} & \texttt{[C2 16]} \\
    \textit{missing} & \texttt{[C2 17]} \textit{drawn}\\
    {}\texttt{[C3 0]} & \texttt{[C3 0]}
\end{tabular}

@@

>>>>

\vspace{\baselineskip}
\begin{tabular}{l l}
    \textbf{old} & \textbf{new} \\ \hline
    {}\texttt{[C1 0]} & \texttt{[C1 0]} \\
    {}\texttt{[C2 0]} & \texttt{[C2 0]} \\
    {}\texttt{[C2 1]} & \textit{unused} \\
    {}\texttt{[C1 16]} & \texttt{[C1 16]} \\
    {}\texttt{[C1 17]} & \texttt{[C1 17]} \\
    {}\texttt{[C1 18]} & \textit{unused} \\
    {}\texttt{[C2 16]} & \texttt{[C2 16]} \\
    \textit{missing} & \texttt{[C2 17]} \textit{drawn}\\
    {}\texttt{[C3 0]} & \texttt{[C3 0]}
\end{tabular}
\vspace{\baselineskip}

@@

\section{Case Study}

The description of Anglican language and environment would be
incomplete without a case study showing the process of building
and executing a solution of an inference problem. This case
study takes a problem for which a solution is not immediately
obvious, \textit{the Deli dilemma}, and guides through
writing an Anglican program for the problem, executing the
program, and post-processing results.

\subsection{The deli dilemma}

Imagine that we are facing the following dilemma:

\begin{quote}
A customer wearing round sunglasses came to a deli at 1:13pm,
and grabbed a sandwitch and a coffee. Later on the same day, a
customer wearing round sunglasses came at 6:09pm and ordered a
dinner.  Was it the same customer?
\end{quote}

In addition to the story above, we know that:

\begin{itemize}
    \item There is an adjacent office quarter, and it takes
        between 5 and 15 minutes on average to walk from an
        office to the deli, were different average times are  for
        different buildings in the quarter.
    \item Depending on traffic lights, the walking time varies by about 2
        minutes.
    \item The lunch break starts at 1:00pm, and the workday ends at 6:00pm.
    \item The waiter's odds that this is the same customer are 2 to 1.
\end{itemize}

\subsection{Anglican query}

We want to formalize the dilemma in an Anglican query, based on
the knowledge we have. Let us formalize the knowledge in
Clojure (the times are in minutes). First, we encode our prior
information which holds true independently of the customer's
visit:

\begin{lstlisting}[style=default]
(def mean-time-to-arrive
     "average time to arrive"
     10.)
(def sd-time-to-arrive
     "standard deviation of arrival time"
     3.)
(def time-sd
     "walking time deviation"
     1.)
\end{lstlisting}

Then, we encode our observations, based on which we want to
solve the dilemma:

\begin{lstlisting}[style=default]
(def lunch-delay
     "time between lunch break and lunch order"
     13.)
(def dinner-delay
     "time between end of day and dinner order"
     9.)
(def p-same
     "prior probability of the same customer"
     (/ 2. 3.))
\end{lstlisting}

For inference, one often chooses known distribution to
represent uncertainty. We choose the normal
distribution for representing uncertainty about average arrival
time.

\begin{lstlisting}[style=default]
(def time-to-arrive-prior 
     "prior distribution of average arrival time"
     (normal mean-time-to-arrive
             sd-time-to-arrive))
\end{lstlisting}

There are two possibilities, either the same customer visited
the deli twice, or two different customers came to the deli,
one for lunch, one for dinner. We define an Anglican function
for each case. Note that this is the first time we switch from
Clojure to Anglican.

\begin{lstlisting}[style=default]
(defm same-customer 
  "observe the same customer twice"
  [time-to-arrive-prior
   lunch-delay
   dinner-delay]
  (let [time-to-arrive
          (sample time-to-arrive-prior)]
    (observe (normal time-to-arrive time-sd) 
             lunch-delay)
    (observe (normal time-to-arrive time-sd)
             dinner-delay)
    (predict :time-to-arrive time-to-arrive)))

(defm different-customers
  "observe different customers"
  [time-to-arrive-prior
   lunch-delay
   dinner-delay]
  (let [time-to-arrive-1
          (sample time-to-arrive-prior)
        time-to-arrive-2
          (sample time-to-arrive-prior)]
    (observe (normal time-to-arrive-1 time-sd) 
             lunch-delay)
    (observe (normal time-to-arrive-2 time-sd)
             dinner-delay)
    (predict :times-to-arrive
      [time-to-arrive-1 time-to-arrive-2])))
\end{lstlisting}

Both functions have the same structure: we first `guess' the
average arrival time and then observe the actual time from a
distribution parameterized by the guessed time. However, in
\texttt{same-customer} the average arrival time is the same for
both the lunch and the dinner, while for
\texttt{different-customers} two average arrival times are
guessed independently.

We are finally ready to define the query.

\begin{lstlisting}[style=default]
(defquery deli [time-to-arrive-prior
                lunch-delay
                dinner-delay]
 (let [is-same-customer (sample (flip p-same))
       observe-customer (if is-same-customer
                           same-customer
                           different-customer)]
    (predict :is-same-customer is-same-customer)
    (observe-customer time-to-arrive-prior
                      lunch-delay
                      dinner-delay)))
\end{lstlisting}

Performing inference on query \texttt{deli} should predict the
probability that the same customer visited the deli twice, as
well as probability distributions of average arrival times for
both cases.

\subsection{Inference}
